---
title: "SVD analyses and data exploration"
format: 
  docx:
    fig-dpi: 300
  # html: default
execute: 
  echo: false
  warning: false
  messages: false
---

```{r setup}
library(see)
# library(performance)
# library(patchwork)
source(here::here("R/functions.R"))
df_all <- targets::tar_read(df_complete)

df <- df_all |> dplyr::filter(include)

df_format <- targets::tar_read(df_formatted)
```

These are the planned analyses, tables and figures for the main article.

Drop all the questions and save them for a later specific publication.

Agreed upon scoring overview:

```{r}
df_format |>
  svd_systems_overview(
    list.system = svd_scoring_systems(),
    subset = match("huijts_simple_ext_svd", names(svd_scoring_systems()))
  ) |>
  # gtsummary::as_gt() |>
  # gt::tab_header(title = "SVD scoring system")|>
  gt_theme()
```

# Baseline table

```{r}
#| tbl-cap: Baseline factors
baseline_table <- df_format |>
  dplyr::mutate(svd_score_04 = simple_score_f) |>
  # get_vars(c("preds")) |>
  dplyr::select(
    dplyr::all_of(purrr::pluck(model_input(), "main_exp")),
    svd_score_04, simple_microbleed, simple_lacunes, simple_wmh, simple_atrophy,
    # nihss,
    pase_0_q
  ) |>
  dplyr::select(
    svd_score_04,
    dplyr::starts_with("simple_"),
    dplyr::everything()
  ) |>
  dplyr::mutate(
    pase_0_q = forcats::fct_rev(pase_0_q),
    svd_score_04 = as.numeric(svd_score_04)
  ) |>
  # dplyr::mutate(female_sex = dplyr::if_else(female_sex, "Female", "Male")) |>
  # dplyr::transmute(simple_score, age, female_sex= dplyr::if_else(female_sex, "Female", "Male"), nihss, tpa, evt, pase_0, alone) |>
  labelling_data() |>
  gtsummary::tbl_summary(
    by = pase_0_q,
    missing = "no", type = list(svd_score_04 = "continuous"),
    label = list(
      simple_microbleed = "Microbleeds (any)",
      simple_lacunes = "Lacunes (any)",
      simple_wmh = "Begininng confluenting areas of WMH (Fazekas 2-3)",
      simple_atrophy = "Moderate-Severe global atrophy (GCA 2-3)"
    )
  ) |>
  gtsummary::add_p() |>
  gtsummary::add_overall() |>
  bstfun::add_variable_grouping(
    "SVD subscore items" = c("simple_microbleed", "simple_lacunes", "simple_wmh", "simple_atrophy")
  )

baseline_table |>
  gtsummary::as_gt() |>
  # gt::tab_header(title = "Baseline characteristics")|>
  gt_theme()

# baseline_table |>
# gtsummary::add_p() |>
# gtsummary::bold_p()

# baseline_table |> save_table("baseline_table.docx")
```

NIHSS fordeling er associeret til PASE, men ikke til SVD score(!)

```{r}
c("huijts_simple_ext_svd") |>
  purrr::map(\(.x){
    name <- subset_named_labels(.x)
    df_format |>
      # dplyr::mutate(age_grp = cut(age,c(0,65,100),labels=c("young","old"))) |>
      svd_system_calc() |>
      vertical_stacked_bars(
        score = .x,
        t.size = 6
      ) +
      ggplot2::theme(legend.position = "none") +
      ggplot2::ggtitle(name) +
      ggplot2::ggtitle("Small Vesel Disease burden score")
  })

c("huijts_simple_ext_svd") |>
  purrr::map(\(.x){
    name <- subset_named_labels(.x)
    df_format |>
      svd_system_calc() |>
      vertical_stacked_bars(score = .x, t.size = 2) +
      ggplot2::theme( # legend.position = "none",
        axis.title = ggplot2::element_text(),
        text = ggplot2::element_text(size = 10)
      ) +
      # ggplot2::ggtitle("") +
      # ggplot2::ggtitle("Smal Vesel Disease burden score") +
      ggplot2::xlab("Physical acitivty level") +
      ggplot2::ylab("Distribution")
  }) |>
  (\(.x){
    ggplot2::ggsave("svd_x_pa_dist.png",
      plot = .x[[1]], width = 140, height = 80, units = "mm", dpi = 600
    )
  })()
```

# Regression analyses

```{r}
#| tbl-cap: Minimaly adjusted models with age and sex.
#| include: false
purrr::pluck(targets::tar_read(lst_multi_olr_orig), "complete_list")[[1]] |>
  fix_labels() |>
  gtsummary::as_gt() |>
  # gt::tab_header(title = "Planned minimally adjusted analyses")|>
  gt_theme()
```

## Independent correlations in multivariable analysis

```{r}
# targets::tar_read(list_multi_var_olr) |>
#   gtsummary::as_gt() |>
#   # gt::tab_header(title = "Ordinal logistic regression analyses")|>
#   gt_theme()
```

```{r}
targets::tar_read(list_multi_var_olr_steps) |>
  gtsummary::as_gt() |>
  # gt::tab_header(title = "Ordinal logistic regression analyses")|>
  gt_theme()
```

```{r}
ms <- targets::tar_read(df_formatted) |>
  svd_system_calc() |>
  dplyr::select(tidyselect::all_of(c(
    purrr::pluck(model_input(), "gen_exp"),
    purrr::pluck(model_input(), "main_exp"),
    "pase_0_q",
    "huijts_simple_ext_svd"
  ))) |>
  dplyr::mutate(
    huijts_simple_ext_svd = factor(huijts_simple_ext_svd) # ,
    # pase_0_q = forcats::fct_rev(pase_0_q)
  ) |>
  models_uni_mini_multi_olr(out = "huijts_simple_ext_svd")

insight::model_info(ms$Univariable)
insight::supported_models()

caret::varImp(ms$Univariable)

car::vif(ms$Multivariable)

performance::model_performance(ms$Multivariable)

broom::glance(ms$Multivariable)

performance::compare_performance(ms)

ms |> lapply(performance::model_performance) |> dplyr::bind_rows()
```

```{r}
ds <- targets::tar_read(df_formatted)
genodds::genodds(ds$simple_score, project.aid::quantile_cut(ds$pase_0, 2, group.names = c("low", "high")),
  # strata = project.aid::quantile_cut(ds$age,groups = 2,group.names = c("y","O"))
  strata = ds$female_sex
)
```

```{r}
targets::tar_read(df_formatted) |>
  (\(.x){
    split(.x, .x$female_sex)
  })() |>
  setNames(c("MALE", "FEMALE")) |>
  lapply(\(ds){
    genodds::genodds(
      ds$simple_score,
      # strata = project.aid::quantile_cut(ds$age,groups = 2,group.names = c("y","O")),
      project.aid::quantile_cut(ds$pase_0, 2, group.names = c("low", "high"))
    )
  })
```



### Complete simple interaction analyses

```{r}
#| include: false
targets::tar_read(list_multi_var_olr_interact) |>
  purrr::map(fix_labels) |>
  (\(.x){
    gtsummary::tbl_stack(.x, glue::glue("{fix_labels_raw(names(.x))} minimal interaction analysis"))
  })() |>
  gtsummary::as_gt() |>
  gt::tab_footnote(footnote = "All analyses adjusted for age and sex") |>
  # gt::tab_header(title = "Ordinal logistic regression analyses")|>
  gt_theme()
```

### Merged table of adjustments and interact

```{r}
#| tbl-cap: Merged table of adjustments and interact
#| include: false
targets::tar_read(lst_olr_interact_merged) |>
  purrr::imap(\(.x, .i) {
    purrr::map(.x, \(.y) {
      gtsummary::tbl_regression(.y, exponentiate = TRUE) |>
        compress_logical_vars()
    }) |>
      tbl_merge_named() |>
      gtsummary::remove_row_type(
        variables = -dplyr::one_of(.i),
        type = "all"
      )
  }) |>
  gtsummary::tbl_stack() |>
  fix_labels() |>
  gtsummary::as_gt() |>
  gt::tab_footnote(footnote = "All analyses adjusted for age and sex") |>
  # gt::tab_header(title = "Ordinal logistic regression analyses")|>
  gt_theme()
```

#### Main analysis by sex

```{r}
sex_strat <- targets::tar_read(df_formatted) |>
  (\(.x){
    split(.x, .x$female_sex)
  })() |>
  setNames(c("MALE", "FEMALE")) |>
  purrr::map(\(.x){
    ls <- planned_multi_olr(.x, gen_exp = "age") |>
      purrr::pluck("complete_list")

    ls[[1]] |>
      fix_labels()
  })
sex_strat |> gtsummary::tbl_merge()
```


```{r}
l.sex <- targets::tar_read(df_formatted) |>
  (\(.x){
    .x |>
      dplyr::mutate(
        pase_0_q = project.aid::quantile_cut(
          x = pase_0,
          groups = 4,
          group.names = paste0("Q", 1:4)
        )
      ) |>
      split(.x$female_sex)
  })() |>
  setNames(c("MALE", "FEMALE")) |>
  lapply(\(.x){
    mulitvar_olr(.x, reg.fun = regression_uni_mini_multi_olr)
  })

l.sex |>
  (\(.x){
    gtsummary::tbl_stack(
      tbls = .x,
      group_header = names(.x)
    )
  })()
```


```{r}
c("huijts_simple_ext_svd") |>
  purrr::map(\(.x){
    name <- subset_named_labels(.x)
    df_format |>
      dplyr::mutate(
        age_grp = cut(age, c(0, 65, 100), labels = c("young", "old")),
        sex = c("male", "female")[female_sex + 1]
      ) |>
      svd_system_calc() |>
      vertical_stacked_bars(
        score = .x,
        strata = "sex",
        t.size = 6
      ) +
      ggplot2::theme(legend.position = "none") +
      ggplot2::ggtitle(name) +
      ggplot2::ggtitle("Small Vesel Disease burden score stratified by sex")
  })
```


#### Distribution with age stratification

```{r}
c("huijts_simple_ext_svd") |>
  purrr::map(\(.x){
    name <- subset_named_labels(.x)
    df_format |>
      dplyr::mutate(age_grp = cut(age, c(0, 65, 100), labels = c("young", "old"))) |>
      svd_system_calc() |>
      vertical_stacked_bars(
        score = .x,
        strata = "age_grp",
        t.size = 6
      ) +
      ggplot2::theme(legend.position = "none") +
      ggplot2::ggtitle(name) +
      ggplot2::ggtitle("Small Vesel Disease burden score")
  })
```


## Collinearity

```{r}
mod <- df_format |>
  svd_system_calc() |>
  dplyr::select(tidyselect::all_of(c(
    purrr::pluck(model_input(), "gen_exp"),
    purrr::pluck(model_input(), "main_exp"),
    "pase_0_q",
    "huijts_simple_ext_svd"
  ))) |>
  dplyr::mutate(
    huijts_simple_ext_svd = factor(huijts_simple_ext_svd) # ,
    # pase_0_q = forcats::fct_rev(pase_0_q)
  ) |>
  multivariable_olr(
    out = "huijts_simple_ext_svd"
  )

performance::check_model(mod)
```

```{r}
# df_inv <- df_format |>
#   svd_system_calc()
#
# pairs(dplyr::select(df_inv,age,huijts_simple_ext_svd,pase_0))
```
